---
title: "GRAPHICS DEVICES"
Name: Murtip Gajjar
Date: Jun 29
output: word_document
---
HOW TO GENERATE A PLOT

FIRST APPROACH

Call a plotting function using either plot, xyplot, qplot

```{r}
with(faithful, plot(eruptions,waiting))
title(main = "Old Faithful Geyser data") 
```

the plot will appear in the screen
Annotate the generated plot with a title

SECOND APPROACH
Open the pdf device and create it in the working directory
create the plot and sendit to the created file
Annotate the plot with title
finally, close the PDF file device


```{r}

pdf(file = "myplot.pdf")  


 with(faithful, plot(eruptions, waiting))  
 

 title(main = "Old Faithful Geyser data")  

 dev.off()  

```

COPY THE PLOTS TO ANOTHER DEVICE
We use the dev.copy() to copy plots from one device to another. while dev.copy2pdf() is used to copy a plot from the current device to PDF file.

```{r}
 library(datasets)

 with(faithful, plot(eruptions, waiting))  

 title(main = "Old Faithful Geyser data")  

 dev.copy(png, file = "geyserplot.png")  

 dev.off() 
```



