---
title: "K-Means Clustering"
Name : Murtip Gajjar 
Date : July 9 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The K-means clustering algorithm is another bread-and-butter algorithm in high-dimensional data analysis that dates back many decades now
The K-means algorithm produces

A final estimate of cluster centroids (i.e. their coordinates)

An assignment of each point to their respective cluster
```{r}
set.seed(1234)
x <- rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y <- rnorm(12, mean = rep(c(1, 2, 1), each = 4), sd = 0.2)
plot(x, y, col = "blue", pch = 19, cex = 2)
text(x + 0.05, y + 0.05, labels = as.character(1:12))
```
 Here we simulate some data from three clusters and plot 
##Using the kmeans() function
The kmeans() function in R implements the K-means algorithm and can be found in the stats package, 
Two key parameters that you have to specify are x, which is a matrix or data frame of data, and centers which is either an integer indicating the number of clusters or a matrix indicating the locations of the initial cluster centroids.
 
```{r}
dataFrame <- data.frame(x, y)
 kmeansObj <- kmeans(dataFrame, centers = 3)
 names(kmeansObj)
```
You can see which cluster each data point got assigned to by looking at the cluster element of the list returned by the kmeans() function.


##Building heatmaps from K-means solutions
A heat map or image plot is sometimes a useful way to visualize matrix or array data. 
 
First, we need to find the K-means solution.
```{r}
set.seed(1234)
dataMatrix <- as.matrix(dataFrame)[sample(1:12), ]
kmeansObj <- kmeans(dataMatrix, centers = 3)
```

Then we can make an image plot using the K-means clusters.


```{r}
par(mfrow = c(1, 2))
image(t(dataMatrix)[, nrow(dataMatrix):1], yaxt = "n", main = "Original Data")
image(t(dataMatrix)[, order(kmeansObj$cluster)], yaxt = "n", main = "Clustered Data")
```
 
The plot above orders the rows of the matrix/image so that all of the rows in the same cluster are grouped together.
